---
title: "Introduction to ggpie"
author: "Rob Kabacoff"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Pie charts are controversial in data science and statistics. Humans are better at judging length than they are volume. For this reason, bar charts are often a more effective method of visualizing differences among percentages. 

However, pie charts are ubiquitous in industry, and when one is trying to visual part-whole relationships, rather than compare frequencies, they can be effective. 

It bothers me that R has no simple and attractive way to create pie charts - thus this package. The graphs returned are `ggplot2` graphs, and can be further manipulated using that system.

## A Single Pie Chart

The `mpg` data frame in the `ggplot2` package contains data on the fuel economy of automobiles from 1999 to 2008. We'll focus on the variables `year` and `class`. To create a single pie chart displaying the percentage of car classes, use code like the following:

```{r}
library(ggplot2)
library(ggpie)
ggpie(mpg, class)
```

By default, pie slices are labeled with percentages and a legend is created.

The position of the labels is controlled with the `offset` parameter. It captures the distance of the label from the origin of the pie. 

* A value of 0.5 is centered in each slice (see above).   
* A value of 1.0 is centered on the radius of the pie.   
* A value greater than 1.0 will place the label outside the pie.  

```{r}
ggpie(mpg, class, offset = 1 )
```


```{r}
ggpie(mpg, class, offset = 1.2 )
```

If the option `legend = FALSE` is included, each slice is labeled directly.

```{r}
ggpie(mpg, class, offset = 1.3 , legend=FALSE)
```

Many addtional options are available. For example

```{r}
ggpie(mpg, class, 
      legend=FALSE, 
      border.color="white", 
      label.color="white", 
      label.size = 5, 
      offset = .7, 
      percent=FALSE, 
      title="Automobile Models by Class") +
  scale_fill_brewer(palette = "Dark2")
```

Note the use of the `ggplot2` color brewer function to change the color scheme above.
See [scale_color_brewer](https://ggplot2.tidyverse.org/reference/scale_brewer.html) for details.

## Mulitiple Pie Charts

You can include a *by* variable in the function. This will create a separate pie chart for each level of the *by* variable.

```{r}
ggpie(mpg, class, year, 
      offset=1.3, 
      label.size=3,
      legend=FALSE, 
      title="Car Class by Year")
```

The graph below demonstrates several additional options, and how `ggplot2` functions can be added to a graph created by the `ggpie` function.

```{r}
ggpie(mpg, drv, trans, 
      nrow=3,                    # number of rows
      border.color="white",      # border color
      border.width=1.5,          # border width
      label.color="black",       # label color 
      label.size=3,              # label size
      offset=.7,                 # label position
      title="Auto Drive Type by Transmission\n") +
  
  # place legend in lower right (x=.8, y=.2, where each is on a 0-1 scale)
  theme(legend.position=c(.8, .2)) + 
  # specify colors manually
  scale_fill_manual(values=c("yellow", "skyblue", "green"))
```

See the `ggpie` function documentation for more details.
